Draft Documents

### 2022-11-17 src/Selectively_remove_variables.Rmd
probably final result

t(coef(ridge_model))
```
Plotting the coefficients against different lambda values.

@@

```{r}
fit <- predict(glm_Ridge, train.x)
true <- train.y
summary_Ridge_train <- eval_results(fit, true)
summary_Ridge_train
```
The training dataset produced a much higher value compared to the regular multiple regression model shown above.

@@

```{r}
fit <- predict(glm_Ridge, test.x)
true <- test.y
summary_Ridge_test <- eval_results(fit, true)
summary_Ridge_test
```
A slightly lower R^2 value, but still proves that the model is good enough.

>>>>

head(t(coef(ridge_model)))
```
Plotting the coefficients against different lambda values.

@@

```{r}
fit <- predict(glm_ridge, train.x)
true <- train.y
summary_Ridge_train <- eval_results(fit, true)
summary_Ridge_train
```
The training dataset produced a much higher value compared to the regular multiple regression model shown above.

@@

```{r}
fit <- predict(glm_ridge, test.x)
true <- test.y
summary_Ridge_test <- eval_results(fit, true)
summary_Ridge_test
```
A slightly lower R^2 value, but still proves that the model is good enough.

### 2022-11-17 src/draft.Rmd
probably final result

---
title: "Draft Documents"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: citations.bib
urlcolor: blue
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = F)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
# inline hook to 2 digits
inline_hook <- function(x) {
    if(is.numeric(x)) x <- round(x, 2)
    paste(as.character(x), collapse=", ")
}
knit_hooks$set(inline = inline_hook)
# code chunk out put to 
options(digits = 5)
```

@@

knitr::kable(summary_3models[1:3,-5], digits = 3)
knitr::kable(summary_3models[4:6,-5], digits = 3)

```
From the 3 models we can see that the original multi linear regression model is the best one with the highest R^squared value. However the Lasso Regression Model has the lowest MSE, MAE and RMSE. We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting as many values will get removed. Next we will build other models to see if they work better.

## 3.2. Selectively remove variables with high missing rate (Callista)

@@

```{r}
#Remove lgdp.dflt
lm_3 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
summary(lm_3)
plot(lm_3, which = 5)
```
```{r}
#remove lfdi
lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
summary(lm_4)
plot(lm_4, which = 5)
```
Our model shows that among all the 14 predictor variables that we started with, only 11 of them showed significance in building the model.

@@

adjR2 <- function(pred, orig, k) {
  R2 <- r2(pred, orig)
  n <- length(pred)
  adjr2 <- 1 - (1-R2)*(n-1)/(n-k-1)
  return (adjr2)
}
```
Adjusted R-Squared on train and test
```{r}
fit1.summ <- lapply(fit1, \(f) {
  # number of variables
  k <- length(f$coefficients) - 1
  # predict value of test
  pred <- predict(f, countries.test.4)
  test.resid <- pred - countries.test.4$pov
  data.frame(
    `Train MSE` = mean(summary(f)$residuals^2),
    `Test MSE` = mean(test.resid^2),
    `Train MAE` = mean(abs(summary(f)$residuals)),
    `Test MAE` = mean(abs(test.resid)),
    `Train Adj.R2` = summary(f)$adj.r.squared,
    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
  )
})

# number of variables
k <- length(fit2.dummy$coefficients) - 1
# number of variables
fit2.pred <- predict(fit2.dummy, countries.test.4)
test.resid <- fit2.pred - countries.test.4$pov
fit2.summ <- data.frame(
  `Train MSE` = mean(summary(fit2.dummy)$residuals^2),
  `Test MSE` = mean(test.resid^2),
  `Train MAE` = mean(abs(summary(fit2.dummy)$residuals)),
  `Test MAE` = mean(abs(test.resid)),
  `Train Adj.R2` = pool.r.squared(fit2, adjusted = T)[,'est'],
  `Test Adj.R2` = adjR2(fit2.pred, countries.test.4$pov, k)
)

@@

# 4. Conclusion
# 5. References

>>>>

---
title: "Draft Documents"
author: "Team 1: Le Van Minh, Callista Stephine Yu, Muhammad Imtiyaaz B M A M , Thum Jun Long Issac"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: citations.bib
urlcolor: blue
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = F)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
# inline hook to 2 digits
inline_hook <- function(x) {
    if(is.numeric(x)) x <- round(x, 2)
    paste(as.character(x), collapse=", ")
}
knit_hooks$set(inline = inline_hook)
# code chunk out put to 
options(digits = 3)
```

# 1. Introduction
The project aim is to explore different models using the same dataset to evaluate which method will be the best to build a model for the dependent variable. The dataset we are using for this project has 1901 data values from 23 different variables. Our plan for this project is to first build a model with all the variables in the dataset. We will then follow it up by builidng a model without the missing values and build a ridge regression and Lasso regression model. Our second approach will be to selectively remove variables from the model and to build the best model  using regularisation and regression. Our next approach was to add the variables into the model one at a time to see if a different model can be achieved. Lastly, we performed imputation on our model to tackle the missing values present in our dataset.We then compared all of this approaches to come to a conclusion.

Github repo: [https://github.com/IntrovertHedgehog/PublicExpenditureAndPoverty](https://github.com/IntrovertHedgehog/PublicExpenditureAndPoverty)

@@

knitr::kable(summary_3models[1:3,-5], digits = 3)
knitr::kable(summary_3models[4:6,-5], digits = 3)
```
From the 3 models we can see that the original multi linear regression model is the best one with the highest R^squared value. However the Lasso Regression Model has the lowest MSE, MAE and RMSE. We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting as many values will get removed. Next we will build other models to see if they work better.

## 3.2. Selectively remove variables with high missing rate

@@

```{r}
#Remove lgdp.dflt
lm_3 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
summary(lm_3)
plot(lm_3, which = 5)
```

```{r}
#remove lfdi
lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
summary(lm_4)
plot(lm_4, which = 5)
```
Our model shows that among all the 14 predictor variables that we started with, only 11 of them showed significance in building the model.

@@

adjR2 <- function(pred, orig, k) {
  R2 <- r2(pred, orig)
  n <- length(pred)
  adjr2 <- 1 - (1-R2)*(n-1)/(n-k-1)
  return (adjr2)
}
```
Adjusted R-Squared on train and test
```{r}
fit1.summ <- lapply(fit1, \(f) {
  # number of variables
  k <- length(f$coefficients) - 1
  # predict value of test
  pred <- predict(f, countries.test.4)
  test.resid <- pred - countries.test.4$pov
  data.frame(
    `Train MSE` = mean(summary(f)$residuals^2),
    `Test MSE` = mean(test.resid^2),
    `Train MAE` = mean(abs(summary(f)$residuals)),
    `Test MAE` = mean(abs(test.resid)),
    `Train R2` = summary(f)$r.squared,
    `Test R2` = r2(pred, countries.test.4$pov)
  )
})

# number of variables
k <- length(fit2.dummy$coefficients) - 1
# number of variables
fit2.pred <- predict(fit2.dummy, countries.test.4)
test.resid <- fit2.pred - countries.test.4$pov
fit2.summ <- data.frame(
  `Train MSE` = mean(summary(fit2.dummy)$residuals^2),
  `Test MSE` = mean(test.resid^2),
  `Train MAE` = mean(abs(summary(fit2.dummy)$residuals)),
  `Test MAE` = mean(abs(test.resid)),
  `Train R2` = pool.r.squared(fit2)[,'est'],
  `Test R2` = r2(fit2.pred, countries.test.4$pov)
)

@@

# 4. Conclusion

Approach 1, 2 and 4 seem to have the best performace.

First Approach:
```{r}
knitr::kable(summary_3models[1:3,-5], digits = 3)
knitr::kable(summary_3models[4:6,-5], digits = 3)
```

Second Approach:

```{r}
knitr::kable(summary_lasso, digits = 3)
knitr::kable(summary_Ridge, digits = 3)
```

Fourth Approach:
```{r}
kable(result.df[c('alpha', 'lambda', 'Train.R2', 'Test.R2')])
```

We can deduce our final interpretation from these models.

```{r}
var <- c('incomeL', 'incomeLM', 'incomeUM', 'edu.total', 'hlth', 'mil')
result1 <- coefficients(model3)[var]
# result1[names(scaler2)] <- result1[names(scaler2)] / scaler2[names(scaler2)]
# result1 <- result1 * scaler2['pov']
# result1 <- result1[var]

result2 <- coefficients(glm_Ridge)[,'s0']
result2[names(scaler2)] <- result2[names(scaler2)] / scaler2[names(scaler2)]
result2 <- result2 * scaler2['pov']
result2 <- result2[var]

result4 <- res[2,var]

# combine
finalResults <- data.frame(Approach = c(1,2,4))
finalResults <- finalResults %>% 
  cbind(rbind(result1, result2, result4))

kable(finalResults)
```

Some important interpretation:

- Low income countries generally have higher poverty index than High income countries.
- As suggested by 1st approach, Lower-Middle income countries have higher poverty index than High income countries. While 2nd and 4th approach suggest otherwise.
- Upper-Middle income countries in general have lower poverty index than High income countries.

As suggested by these models. Upper-Middle income countries have lowest poverty rate. This can be attributed to that these countries tend to have more even wealth distribution than high income countries, and have enough money to alleviate poor people.

- Spending on education is negatively correlated with poverty. On average, 1% (of GDP) more spending in education results in 3 point decrease in poverty index. This suggests that public spending on education can reduce poverty.
- Spending on Healthcare is negatively correlated with poverty. On average, 1% (of GDP) more spending in education results in 0.5 point decrease in poverty index.
- Spending on military has ambiguous effect on poverty. Different model results in different interpretation. This can be accounted to the difference in the data used in each approach. Approach 1 only use complete cases, which caused the model to predict positive relationship. However, as more data is utilised in approach 2&4 use more data, the effect tend to be neutral/negative.

# 5. References
